---
title: "02: Sampling"
author: "[List your group here]"
---

Press the "play" button below to run the chunk of code. This (1) loads the libraries that we need and (2) tells R to read `mission_data.csv` from my website folder into your R session and call it `mission_data`. 

```{r}
library(tidyverse)
#mission_data = read_csv("https://www.math.carleton.edu/aluby/stat120/mission_data.csv")
mission_data = read_csv("../data/mission_data.csv") # update to server data!!
```

Check your "environment" pane in the upper right to make sure you can see a dataset called `mission_data`. Try clicking it, or running `View(mission_data)` in the console to bring up the data viewer. 

**spoiler alert:** The next chunk of code computes the *population mean*. 

```{r}
mean(mission_data$length)
```

```{r}
set.seed(091824) # Sets the random seed so we all get the same answer
sample = sample(1:nrow(mission_data), size = 10) # selects a random sample of size 10 from the numbers 1-364
sample # prints the sample
```

The next chunk of code `slices` our population to draw our sample. Note that the `position` variable should match the `sample` output above. 

```{r}
mission_sample = mission_data %>% 
  slice(sample)

mission_sample
```

```{r}
mean(mission_sample$length)
```

To try other random samples, change (or remove!) the `set.seed()` line of code, and try re-running the rest of the code. Do you ever get a sample mean that looks like your "by hand" sample mean?

When you're done, **knit this file** and try uploading the PDF to gradescope. There are two questions, one for the *population mean* and one for a *sample mean*. Be sure to mark the pages so I can see your answers! 
