---
title: "04: Categorical Variables + Intro to Quantitative Variables"
---

The lab manual data `Pew` contains data from a January 2014 Pew Research Center survey about the Internet. We'll consider two variables from this survey:

- `income`: the person's yearly income, grouped into four categories
- `values`: an answer to the question asking if the Internet has been a "good thing"

```{r, message = FALSE}
library(tidyverse)
```

**0.** Import the data into R by running the following code chunk. Click on the dataset in the environment pane to pull up the data viewer. Can you see the "spreadsheet view"?

```{r}
pew = read_csv("http://www.math.carleton.edu/Stat120/RLabManual/Pew.csv")
```

**1.** The code below computes the `frequency table` for the `values` variable. Add a "pipe" (`%>%`) and another command to answer: what proportion of respondents said that the Internet *has* been a good thing?

```{r}
table(pew$values) 
```

**2.** Using esquisse, make a bar plot of the counts of people who said the Internet has vs. has not been a good thing. Insert your code in the chunk below.

```{r}

```


**3.** Show the two-way table for the `values` and `income` variables. How many people from the survey had an income above $150,000 a year and thought the Internet was a bad thing?

```{r}
table(pew$income, pew$values)
```

**4.** The table in part 3 shows the income categories in the wrong order (i.e., not in order of increasing income). You can fix this by changing the order of the categories of the `income` variable as below (see Appendix A.1 of the lab manual). Remake the table after reordering the categories.

```{r}
pew$income <- factor(pew$income, levels = c("0-30000", "30000-75000",
                                            "75000-150000", "150000"))
```

**5.** Use `esquisse` to make a *stacked barplot* of `income` and `values`. 

```{r}

```

**6.** Your bar plot by default should show the *counts* of each combination. Copy and paste your bar plot code, and change `geom_bar()` to `geom_bar(position = "fill")`. Explain what this chunk of code did. 

```{r}

```

**Stop Here** and let Amanda know you've finished with the categorical EDA section

A third variable in the dataset is `age`. The next few questions walk through how to do basic EDA for a quantitative variable.

**7.** Create a histogram of the `age` variable using esquisse

```{r}

```

**8.** In the empty code chunk below, run `mean(pew$age)` and `median(pew$age)`. Now try `mode(pew$age)`. What happened?

```{r}

```

**9.** Using your answers to the previous two questions, describe the distribution of `age`. 


