---
title: "06: Correlation and Intro to Regression"
---

This dataset gives education-related data for the 50 states and the District of Columbia. The variables are: 

 - `region`: West, Northeast, Midwest, South
 - `pop`: Population, in 1,000's
 - `verbal` and `math`: average SAT verbal and math scores
 - `taken`: percent of students taking the SAT
 - `noHS`: percent of population with no high school diploma
 - `teachersPay`: median teacher salary, in 1,000's

**0.** Load code libraries and the data and make sure you can view it. What is each case? 

```{r}
library(tidyverse)
sat <- read.csv("http://math.carleton.edu/Stat120/RLabManual/sat.csv")
```

We're going to investigate the relationship between math SAT scores (`math`) and the percentage of high school students who took the SAT. 

**1.** *Before looking at the data*, do expect there to be a positive, negative, or *no* relationship? Why? 

**2.** Make a scatterplot of `math` on the y-axis and `taken` on the x-axis with the line of best fit included (see notes from today for the line of code to include). What do you notice?

**3.** Use the `lm()` command to find the equation for this line. *Be careful about the X and Y variables!* Interpret the slope and intercept in context.

**4.** Find the correlation for this relationship. 

**5.** Color the scatterplot by `region`. What do you notice? 

The code chunk below creates a new dataset called `sat_northeast` which filters to only the Northeast states (`==` is code for "equals"). 

```{r}
sat_northeast = sat %>%
  filter(region == "Northeast")
```

**6.** Make a scatterplot of `math` on the y-axis and `taken` on the x-axis with the line of best fit included. 

**7.** Use the `lm()` command to find the equation for this line. How does it compare to your line from (3)? 

**8.** Find the correlation for this relationship. How does it compare to the correlation for the whole dataset? 

*Note:* this data is adapted from Ch 3.4 of the Lab Manual. You can find most of the code solutions there if your group gets stuck! 