---
title: "15: Hypothesis Tests"
---

```{r, message = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
library(CarletonStats)
```

Do heart attack victims have higher cholesterol levels than non-heart attack victims? The dataset `Cholesterol.csv` has cholesterol measurements (mg/dL) for a sample of heart attack victims (4 days after the heart attack) and a sample of non-heart attack victims. 

```{r}
Cholesterol <- read.csv("http://math.carleton.edu/Stat120/RLabManual/Cholesterol.csv")
```

**0.** Load the data and check the "spreadsheet view". What is each case? How many variables are there?

**1.** Create an appropriate graph to answer the research question. 

**2.** *Without performing a formal statistical test* Does there appear to be evidence of a difference in cholesterol levels between the control group and the heart attack group? 

**3.** Write out appropriate null and alternative hypotheses using the original research question.

**4.** Use the `permTest()` command in R to test your hypotheses. (Make sure to include `seed()`!) 

**5.** Report the *test statistic* and *p-value* from the R output. 

**6.** Make a *formal statistical decision* and *report your conclusion in context*.  

**7.** What type of error could you have made in (6)? Do you know the probability that an error occurred?

*When you're done:* Let Amanda know. 

*Note:* This is based on Lab Manual Ch5 #7

**8.** True or False? If false, explain why or correct the statement. 

(a) If a p-value is 10%, there is a 1 in 10 chance the null hypothesis is correct
(b) When a p-value is extremely small, the result is extremely important
(c) A small p-value means that the result could not possibly have been due to chance
(d) A big p-value means that you do not have strong evidence against the null hypothesis
(e) A p-value is the probability of getting a result more in favor of the alternative hypothesis than the result that you observed, assuming the null hypothesis is true
(f) A p-value is the probability of getting a result more in favor of the alternative hypothesis than the result that you observed, assuming the alternative hypothesis is true

