---
title: "04: EDA for Categorical Variables"
format: 
  revealjs: 
    theme: slides.scss
    scrollable: true
    incremental: true
    title-slide-attributes: 
      data-background-color: "#1F4257"
brand: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(xaringanthemer)
library(patchwork)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, fig.align = 'center', fig.width = 10, fig.height = 6)
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
```



## Today: 

1. Wrap up presentations from Friday
1. EDA for 1 Categorical Variable
2. EDA for 2 Categorical Variables
3. Intro to EDA for Quantitative Variables

# Remaining presentations from Friday

# Penguins! 

## Load Data

```{r, echo = TRUE}
penguins = palmerpenguins::penguins
penguins
```

## Frequency Table

```{r, echo = TRUE}
table(penguins$species)
```

## Barplot

```{r, echo = TRUE}
ggplot(penguins, aes(x = species)) + 
  geom_bar()
```

## Barplot: y-axis

```{r, echo = TRUE}
ggplot(penguins, aes(y = species)) + 
  geom_bar()
```

## Barplot: with color! 

```{r, echo = TRUE}
ggplot(penguins, aes(y = species, fill = species)) + 
  geom_bar()
```

## Relationship between `species` and `island`

```{r, echo = TRUE}
table(penguins$species, penguins$island)
```
. . . 

- How would you represent this information using a barplot?

## Relationship between `species` and `island`

```{r, echo = TRUE}
ggplot(penguins, aes(y = species, fill = island)) + 
  geom_bar()
```

## Joint "Distribution"

Distribution = proportions

```{r, echo = TRUE}
table(penguins$species, penguins$island) %>%
  proportions() 
```

## Conditional Distribution

Conditional distribution of `species` on `Dream` island

```{r, echo = TRUE}
table(penguins$species, penguins$island) 
```
. . . 

```{r, echo = TRUE}
table(penguins$species, penguins$island) %>%
  proportions(2) # condition on COLUMNS
```

## Marginal Distribution

Marginal distribution of `species` 

```{r, echo = TRUE}
table(penguins$species, penguins$island) %>%
  addmargins()
```

. . . 

```{r, echo = TRUE}
table(penguins$species) %>%
  proportions()
```



