---
title: "Midterm 1 + Check-in"
format: 
  revealjs: 
    theme: slides.scss
    scrollable: true
    incremental: true
    title-slide-attributes: 
      data-background-color: "#1F4257"
brand: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(xaringanthemer)
library(patchwork)
library(palmerpenguins)
library(googlesheets4)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, fig.align = 'center', fig.width = 10, fig.height = 6)
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
slides_theme = theme_minimal(
  base_family = "Source Sans 3",
  base_size = 20)
theme_set(slides_theme)
```


# Midterm 1 {.maize}


## A note on regrade requests: 

- Grading is often a tedious task, and the grading team will sometimes make mistakes. I am always happy to fix these mistakes, and gradescope makes it easy to do so.
- Regrade requests must be submitted on gradescope within two weekdays after homework has been returned to you. 
- Regrade requests are for administrative errors or obvious grading mistakes. I will not consider regrade requests for anything that applied to the entire class (e.g. “I think this mistake should only be worth 1 point instead of 2” or “I didn’t realize we had to do X”). 
- If you submit two or more inappropriate regrade requests, I will not consider additional regrade requests from you for the remainder of the term. 
- If you’re unsure whether you should file a regrade request or not, just ask! You are always welcome to discuss any grading questions with me in office hours.

# Check-In Results {.maize}

## 

:::: {.columns}

::: {.column width="50%"}
**Going Well**

 - Prep before class (readings, video, pre-class assignment)
 - Note Sheets are helpful
 - Group work
 - Homework
 - Manageable workload
 - Pace is good
 - R is fun/not that bad/etc.
:::

::: {.column width="50%"}
**Things to change:** 

 - R: more code examples on handouts
 - R: explain more step by step
 - Some homework questions are hard to figure out what its asking for
:::

::::

## Balance of lectures/group work

```{r}
survey = read_sheet("https://docs.google.com/spreadsheets/d/1wieyZUUxueThb-Uh45SK15TV7eMQTzUDZ60Gocmx1n4/edit?usp=sharing") |>
  janitor::clean_names()


survey |>
  ggplot(aes(y = how_do_you_feel_about_the_balance_of_lectures_and_group_work)) + 
  geom_bar() +
  labs(
    y = ""
  )
```

## Pre-class videos

```{r}
survey |>
  ggplot(aes(y = how_do_you_feel_about_pre_class_videos)) + 
  geom_bar() +
  labs(
    y = ""
  )
```

## How helpful is....

```{r}
p1 <- survey |>
  ggplot(aes(y = how_helpful_is_going_to_class_for_learning)) + 
  geom_bar() +
  labs(
    y = "",
    title = "Going to class"
  ) + 
  ylim(c(0,10))

p2 <- survey |>
  ggplot(aes(y = how_helpful_is_group_work_for_learning)) + 
  geom_bar() +
  labs(
    y = "",
    title = "Group Work"
  ) + 
  ylim(c(0,10))

p3 <- survey |>
  ggplot(aes(y = how_helpful_are_homework_assignments_for_learning)) + 
  geom_bar() +
  labs(
    y = "",
    title = "Homework"
  ) + 
  ylim(c(0,10))

p1 + p2 + p3
```

## How much time/How often

```{r}
p1 <- survey |>
  mutate( 
     how_much_cat = fct_relevel(how_much_time_outside_of_class_are_you_spending_on_the_course_per_week_this_includes_pre_class_prep_homework_assignments_and_finishing_in_class_r_activities, "Between 6 and 8 hours", after = 1)
   ) |>
  ggplot(aes(y = how_much_cat)) + 
  geom_bar() +
  labs(
    y = "",
    title = "Time outside of class"
  ) 
p2 <- survey |>
  mutate( 
     resource_cat = fct_relevel(how_often_are_you_using_resources_like_stat_lab_office_hours_or_slack, "Once or twice", after = 1)
   ) |>
  ggplot(aes(y = resource_cat)) + 
  geom_bar() +
  labs(
    y = "",
    title = "Resource Frequency"
  ) 

p1 + p2 

```

## My goals for the rest of term: 

- More R code on handouts
- Continue mixing in videos for class prep


