---
title: "Notes 01: Data"
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

# The Structure of Data

::: callout-note
## Cases

\vspace{.5in}

:::


::: callout-note
## Variables

\vspace{.5in}
:::

# Types of Variables

\vspace{2in}


# Examples

For each of the following, label the *cases* and *variables*. For each variable, state whether it is *categorical* or *quantitative*. 

## Penguins

```{r, echo = FALSE}
palmerpenguins::penguins 
```

## Price of diamonds by carat and cut quality

```{r, echo = FALSE}
#| fig-align: left
#| fig-height: 2
#| fig-width: 4

ggplot(diamonds %>% filter(cut %in% c("Good", "Ideal")) %>% sample_n(500), aes(x = carat, y = price, col = cut)) + 
  geom_point(size = .5) + 
  labs(
    x = "Weight (in carats)",
    y = "Price (in USD)",
    color = "Cut Quality"
  )
```

## Is there a "Sprinting Gene"?

A gene called `ACTN3` encodes a protein which functions in fast-twitch muscles. Some people have a variant of this gene that cannot yield this protein. To address the question of whether this gene is associated with sprinting ability, geneticists tested people from three different groups: world-class sprinters, world-class marathon runners, and a control group of non-athletes. In the sames tested, 6% of the sprinters had the gene variant, compared with 18% of non-athletes and 24% of the marathon runners. 

Sketch out a possible dataset, and then answer the questions. 
