---
title: "Notes 04: Categorical Variables"
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

# Categorical Variables

Categorical variables are best summarized with a **frequency table** and visualized using a **barplot**. 


```{r}
#| echo: true
table(penguins$species)
```


```{r}
#| echo: true
#| fig-width: 4
#| fig-height: 2


ggplot(penguins, aes(y = species)) + 
  geom_bar()
```



If we want to summarize a categorical variable with a single number, or choose an appropriate parameter/statistic for a categorical variable, we often use a **proportion**.

::: callout-note
Proportion \vspace{.5in}
:::

When we have two categorical variables, we often use a **two-way table** to summarize them at the same time (also called the **joint distribution**). 


```{r}
#| echo: true
table(penguins$species, penguins$island)
```



We might also care about the **marginal distribution** (the margins) 

```{r, echo = TRUE}
table(penguins$species, penguins$island) |>
  addmargins()
```

or **conditional distribution** (a specific row/column).

```{r, echo = TRUE}
table(penguins$species, penguins$island) |>
  proportions(2) # condition on COLUMNS
```

::: callout-tip
## Independence

Two categorical variables are **independent** if the marginal and conditional distributions are the same. This means that knowing something about one variable tells us nothing about the distribution of the other variable. 
:::

**Example:** Is penguin species independent of island?
\vspace{1in}


Visually, we use a filled barplot: 

```{r}
#| echo: true
#| fig-width: 4
#| fig-height: 2

ggplot(penguins, aes(y = species, fill = island)) + 
  geom_bar()
```

\pagebreak 

**Example:** Below is the two-way table for our class representing the answers to "What is your class year?" and "Coffee or tea?"

|        | First year| Sophomore|
|:-------|----------:|---------:|
|Coffee  |         10|         8|
|Neither |          3|         0|
|Tea     |          5|         5|

a. What is the marginal distribution of *coffee_or_tea*? 

\vspace{1in}

b. What is the conditional distribution of *class year* among those who *prefer coffee*?

\vspace{1in}


c. What is the conditional distribution of *coffee or tea* among those who *sophomores*? 

\vspace{1in}

d. Does *coffee or tea* appear to be independent of *class year*? 



