---
title: "Notes 06: Two Quantitative Variables"
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(palmerpenguins)
options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

When we're interested in the relationship between two quantitative variables, the best visualization is a **scatterplot**. 

```{r}
#| fig-width: 3
#| fig-height: 2
#| fig-align: "center"

ggplot(penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(size = .5) +
  labs(
    x = "Explanatory Variable",
    y = "Response Variable"
  )
```


If we want to summarize the relationship in a single number, we'll often choose the **correlation**.

::: callout-note
Correlation: measure of the strength and direction of a linear relationship between two quantitative variables

- sample: \hspace{1.5in} population: 
- between -1 and +1
- no units, doesn't matter what x and y are
$r = \frac{1}{n-1} \sum (\frac{(x_i - \bar{x})}{s_x})(\frac{(y_i - \bar{y})}{s_y})$
:::

```{r}
#| fig-width: 8
#| fig-height: 2
set.seed(092324)

sim_data = tibble(
  x1 = rnorm(50), 
  y1 = 2*x1,
  x2 = rnorm(50), 
  y2 = -.5*x2 + runif(50),
   x3 = runif(50), 
  y3 = runif(50),
  x4 = rnorm(50), 
  y4 = x4 + rnorm(50, sd = 2)
) 

p1 = ggplot(sim_data, aes(x = x1, y = y1 )) +
  geom_point() +
  labs(
    x = "X",
    y = "Y",
    title = "r = 1"
  )

p2 = ggplot(sim_data, aes(x = x2, y = y2 )) +
  geom_point() +
  labs(
    x = "X",
    y = "Y",
    title = "r = -.86"
  )

p3 = ggplot(sim_data, aes(x = x3, y = y3 )) +
  geom_point() +
  labs(
    x = "X",
    y = "Y",
    title = "r = .14"
  )

p4 = ggplot(sim_data, aes(x = x4, y = y4 )) +
  geom_point() +
  labs(
    x = "X",
    y = "Y",
    title = "r = .52"
  )

(p1 | p2 | p3 | p4)
```

```{r, echo = TRUE}
cor(penguins$body_mass_g, penguins$flipper_length_mm, use = "complete.obs")
```

If the relationship is **linear**, we can also summarize the relationship with "the line of best fit" or "least squares" line. 

```{r, echo = TRUE}
#| fig-width: 3
#| fig-height: 2
#| fig-align: "right"

ggplot(penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(size = .5) +
  geom_smooth(method = "lm", se = FALSE)
```


```{r, results = 'hide', echo = TRUE}
lm(flipper_length_mm ~ body_mass_g, data = penguins) 
```
```{r}
lm(flipper_length_mm ~ body_mass_g, data = penguins) %>% coef()
```

**Interpretation:**

\vspace{2in}

**Two types of outliers:** 

```{r}
#| fig-width: 6
#| fig-height: 2
set.seed(092524)
data = tibble(
  x = c(rnorm(100, 85, 5), 110),
  y = .9*x + rnorm(100, sd = 5) 
) 

data$y[101] = 80

data_filtered = data[-101,]

p1 = data %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = .5) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "With Outlier"
  ) + 
  xlim(c(70, 110))

p2 = data_filtered %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = .5) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Without Outlier"
  ) + 
  xlim(c(70, 110))

p1 + p2

eq1 = lm(y ~ x, data = data)
eq2 = lm(y ~ x, data = data_filtered)
```

```{r, include = FALSE}
equatiomatic::extract_eq(eq1, use_coefs = TRUE)
equatiomatic::extract_eq(eq2, use_coefs = TRUE)
```

Best fit line with outlier: $\operatorname{\widehat{y}} = 7.1 + 0.82\operatorname{x}$ 

Best fit line without outlier: $\operatorname{\widehat{y}} = -8.79 + 1.01\operatorname{x}$

```{r}
#| fig-width: 6
#| fig-height: 2
set.seed(092524)
data = tibble(
  x = c(rnorm(100, 85, 5), 85),
  y = .9*x + rnorm(100, sd = 5) 
) 

data$y[101] = 105

data_filtered = data[-101,]

p1 = data %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = .5) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "With Outlier"
  ) + 
  xlim(c(70, 110))

p2 = data_filtered %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = .5) +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Without Outlier"
  ) + 
  xlim(c(70, 110))

p1 + p2

eq1 = lm(y ~ x, data = data)
eq2 = lm(y ~ x, data = data_filtered)
```

```{r, include = FALSE}
equatiomatic::extract_eq(eq1, use_coefs = TRUE)
equatiomatic::extract_eq(eq2, use_coefs = TRUE)
```

Best fit line with outlier: $\operatorname{\widehat{y}} = -8.1 + 1.01\operatorname{x}$ 

Best fit line without outlier: $\operatorname{\widehat{y}} = -8.79 + 1.01\operatorname{x}$
