---
title: "Notes 10 Addendum: Fingerprints Sampling Distributions"
--- 

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(bookdown)
library(patchwork)
library(palmerpenguins)
library(broom)
library(infer)
library(ggformula)

options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

**Non-random samples:** (from our class)

```{r}
#| fig-width: 8
#| fig-height: 2.5

phats_5 <- c(0.4, 0, 0.2, 0, 0, 1, 0, 0, 0.4, 0, 0.4, 0.2, 0, 0, 0.2, 0.2, 0, 0, 0, 0, 0.2, 0, 0, 0.2, 0, 0.2, 0, 0, 0, 0, 0.8)


phats_10 <- c(0.3, 0.2, 0.4, 0.1, 0, 0.6, 0.3, 0, 0.4, 0, 0.4, 0.3, 0, 0, 0.2, 0.3, 0.1, 0, 0, 0, 0.1, 0, 0, 0.1, 0, 0.1, 0, 0.2, 0.2, 0, 0.7)



p1 <- gf_histogram(~phats_5, bins = 10, col = "white", fill = "black", xlab = "n=5", alpha = .7, boundary = 0) +  scale_x_continuous(breaks = seq(0,1,by=.2), limits = c(0,1)) + ylim(c(0,20))

p2 <- gf_histogram(~phats_10, bins = 12, col = "white", fill = "black", xlab = "n=10", alpha = .7, boundary = 0) + scale_x_continuous(breaks = seq(0,1,by=.2), limits = c(0,1)) + ylim(c(0,20))

p1 + p2
```

- As n increases, there is less spread in the sampling distribution
- Because our samples are biased (our classroom is not a random sample of the population), there's no guarantee that our distribution is getting closer to the "right answer" ($p_{\text{whorl}}$)

**Random samples:** (from fingerprint generator)

```{r}
#| fig-width: 8
#| fig-height: 2.5

phats_5 <- c(0.4, 0.2, 0.2, 0.2, 0.2, 0.2, 0.6, 0.2, 0.6, 0.2, 0.4, 0.4, 0.4, 0.2, 0.2, 0, 0.2, 0.6, 0.4, 0.2, 0.2, 0.6, 0.1, 0, 0.8, 0.2, 0.2, 0.4, 0, 0.2, 0.8, 0.6, 0.2, 0.2, 0.6, 0.6, 0.4, 0.6, 0.2, 0, 0.2, 0.2, 0.2, 0.2, 0.2, 0.4, 0.4, 0, 0.4, 0.2, 0.2, 0.2, 0.2, 0.2, 0.2, 0, 0.4, 0, 0.4, 0.2, 0.2, 0.8, 0.6, 0, 0.4, 0.4, 0.4, 0.2, 0.2, 0.4, 0.2, 0, 0.4, 0, 0.4, 0.2, 0.2, 0.4, 0.4, 0.2, 0.4, 0.2, 0.6, 0.2, 0.8, 0.2, 0.8, 0.1, 0.4, 0, 0.1, 0, 0.6)

phats_20 <- c(0.4, 0.3, 0.45, 0.45, 0.25, 0.25, 0.35, 0.35, 0.35, 0.4, 0.4, 0.2, 0.15, 0.15, 0.15, 0.45, 0.45, 0.45, 0.25, 0.3, 0.3, 0.25, 0.1, 0.2, 0.35, 0.3, 0.25, 0.3, 0.7, 0.35, 0.6, 0.2, 0.3,  0.45, 0.3, 0.3, 0.2, 0.3, 0.25, 0.35, 0.4, 0.05, 0.45, 0.3, 0.3, 0.15, 0.3, 0.45, 0.15, 0.25, 0.45, 0.35, 0.25, 0.35, 0.2, 0.4, 0.3, 0.25, 0.25, 0.7, 0.4, 0.5, 0.4, 0.4, 0.35, 0.55, 0.25, 0.3, 0.3, 0.3, 0.15, 0.25, 0.5, 0.3, 0.4, 0.2, 0.5, 0.15, 0.25, 0.15, 0.4, 0.25, 0.4, 0.35, 0.4, 0.3, 0.35, 0.3, 0.25, 0.45, 0.8, 0.25, 0.6)


p1 <- gf_histogram(~phats_5, bins = 10, col = "white", fill = "black", xlab = "n=5", alpha = .7, boundary = 0) +  scale_x_continuous(breaks = seq(0,1,by=.2), limits = c(0,1)) +  ylim(c(0,40))

p2 <- gf_histogram(~phats_20, bins = 12, col = "white", fill = "black", xlab = "n=20", alpha = .7, boundary = 0) + scale_x_continuous(breaks = seq(0,1,by=.2), limits = c(0,1)) + ylim(c(0,40))

p1 + p2
```

- As n increases, the sampling distribution becomes more symmetric and less spread out
- As n increases, the "peak" of our distribution gets closer and closer to the true parameter
- Shows us the expected **sampling variability** of the sample statistic for a given sample size
