---
title: "Notes 15: More on Testing"
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(bookdown)
library(infer)
library(patchwork)
library(palmerpenguins)
library(broom)

options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

## Types of Errors

|                     | $H_0$ True | $H_0$ False |
|---------------------|:----------:|:-----------:|
| Reject $H_0$        |            |             |
| Do not reject $H_0$ |            |             |

::: callout-note
Type I Error

\vspace{.5in}
:::

::: callout-note
Type II Error

\vspace{.5in}
:::

::: callout-note
$\alpha$

\vspace{.5in}
:::

::: callout-note
$\beta$

\vspace{.5in}
:::

::: callout-note
Power

\vspace{.5in}
:::

**Example:** A public health researcher believes that there is a positive relationship between heart rate and age among ICU patients. Data from 23 patients gives $r = 0.037$. Do you think the probability of a Type II error is higher if $\rho = .9$ or $\rho = .1$? 

$H_0: \rho = 0$

$H_A: \rho > 0$




\pagebreak
**Big Picture** Picture 
\vspace{.5in}

:::: {layout="[0.33, 0.33, 0.33]"}

:::{#firstcol}
**Sampling Distribution**
:::

:::{#secondcol}
**Bootstrap Distribution**
:::

:::{#secondcol}
**Null Distribution**
:::

::::

\vspace{2in}

## Connection to Confidence Intervals

If the value of $H_0$ is *outside* of the confidence interval, we ___________________________________. 

If the value of $H_0$ is *inside* the confidence interval, we ___________________________________. 


**Illustration:** (Social Media Poll Example)

```{r}
#| fig-height: 1.5
#| fig-width: 3
#| layout-nrow: 2
social_media = tibble(
  app = c(rep("twitter", 346), rep("instagram", 530)), 
  visits = c(rep(1, 249), rep(0, 346-249), rep(1, 424), rep(0, 530-424))
)
library(CarletonStats)
test = permTest(visits ~ app, data = social_media)
ci = boot(visits ~ app, data = social_media)

plot(test) + xlim(c(-.1,.2)) + theme(legend.position = "none") + labs(x = "", y = "")
plot(ci) + xlim(c(-.1,.2)) + theme(legend.position = "none") + labs(x = "", y = "")
```

## Statistical vs Practical Significance
