---
title: "Notes 18: The Normal Distribution"
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(bookdown)
library(infer)
library(patchwork)
library(palmerpenguins)
library(broom)

options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

Three main topics of Stat120: 

1. ______________________: Summarizing data with numbers and graphs

2. ______________________: Using confidence intervals to estimate parameters with uncertainty

3. ______________________: Using p-values to evaluate competing hypotheses

Up until now, we've relied on computer simulations (via StatKey or R) to generate __________________ or __________________ distributions. 

We're now going to begin using ________________________________ to generate these distributions instead

::: callout-note
**Normal Density Function:** 

$$f(x) = \frac{1}{\sqrt{2\pi\sigma}} e^{-\frac{1}{2\sigma^2}(x-\mu)^2}$$

Notation:

$~$

Area under the curve = 

$~$

:::

**Example 1:** Verbal SAT scores follow a normal distribution with a population mean of $\mu = 580$ and population standard deviation $\sigma = 70$. What proportion of test-takers score above 650? 

```{r}
#| fig-width: 3
#| fig-height: 2
ggplot(data = data.frame(x = c(580-4*70, 580+4*70)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 580, sd = 70)) +
  scale_y_continuous(breaks = NULL) + 
  scale_x_continuous(breaks=c(580 - 2*70, 580 - 70, 580, 580 + 70, 580+2*70), labels = NULL) +
  labs(y = "density", 
       x = "")
```


\pagebreak

**Example 2:** What is the SAT score for the 90th percentile?

```{r}
#| fig-width: 3
#| fig-height: 2
ggplot(data = data.frame(x = c(580-4*70, 580+4*70)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 580, sd = 70)) +
  scale_y_continuous(breaks = NULL) + 
  scale_x_continuous(breaks=c(580 - 2*70, 580 - 70, 580, 580 + 70, 580+2*70), labels = NULL) +
  labs(y = "density", 
       x = "")
```

::: callout-note
**Z-scores:**

\vspace{.5in}
:::


::: callout-note
**Standard Normal Model**

\vspace{.5in}

:::

::: callout-note
**Central Limit Theorem:** For random samples, if _____ is big enough, the sampling distribution of _______ is approximately ___________________, *regardless of what shape the population distribution is.*
:::

CLT shortcut for testing: 

\vspace{.5in}

CLT shortcut for confidence intervals: 

\vspace{.5in}

Summary of R commands (sketch normal curves to help you remember)

```{r}
#| echo: true
pnorm(650, mean = 580, sd = 70, lower.tail = FALSE)
qnorm(.9, mean = 0, sd = 1, lower.tail = TRUE)
```
