---
title: "Notes 19: CLT-Based Inference for Proportions"
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(bookdown)
library(infer)
library(patchwork)
library(palmerpenguins)
library(broom)

options(digits = 4)
notes_theme = theme_minimal(base_size = 12, base_family = "Source Sans 3")
theme_set(notes_theme)
```

**CLT:** The Central Limit Theorem (CLT) tells us that if the sample size is big enough and the sample is random, 

$$ \bar{X} \sim N(\rule{1cm}{0.15mm}, \rule{1cm}{0.15mm}) $$

$$ \hat{p}\sim N(\rule{1cm}{0.15mm}, \rule{1cm}{0.15mm}) $$

The general form for a **confidence interval** is: 

\vspace{1in}

**Example:** Finding $z^*$

- 95% confidence interval 
- 68% confidence interval
- 99% confidence interval
- 90% confidence interval

## How big is big enough? 

**Example 1:** $\hat{p} = .5$

\vspace{.5in}

**Example 1:** $\hat{p} = .05$

\vspace{.5in}

::: callout-note
**Rule of Thumb for Proportions**: 

- Expected count in each category (Yes/No) should be > ____
- $np > $ ____ and $n(1-p) > $ ____

:::

## How do we find the SE?

::: callout-note
**Standard Error for Proportions**: 

\vspace{1in}

:::

**Idea:** As $n$ gets bigger, the SE gets ________________. If $\hat p$ is close to .5, the SE is  ________________ than if $\hat p$ is close to 0 or 1

**Example:** ESP Example Again

n = 14

```{r}
#| echo: true

p_hat = 3/14
n = 14
SE_p = sqrt((p_hat*(1-p_hat))/n)
z_score = (p_hat - .2)/SE_p
p_val = pnorm(z_score, lower.tail = FALSE)
p_val
```

n=1400:


n=14000:

**Big Picture** Picture


